---
title: "Week 8: HDBSCAN Clustering P3, Merging with AQS" 
subtitle: "SURA 2022"
author: "Shih-Ting (Cindy) Huang"
date: '2022-07-29'
output: html_document
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo=FALSE)

library(tidyverse)
library(dplyr)
library(leaflet)
library(sf)
library(mapview)
library(ggplot2)
library(RColorBrewer)
library(dbscan)
library(leafpop)
library(fpc)
library(lwgeom)
```


```{r load-data, echo=FALSE, include=FALSE}
# Run the script to preprocess fire datasets to include information on time, location, and FRP
url1 <- "https://raw.githubusercontent.com/cindyellow/SURA-2022/main/preprocess_fire_kml.R"
if (!file.exists("../preprocess_fire_kml.R"))
  download.file(url1, destfile = "../preprocess_fire_kml.R")
source("../preprocess_fire_txt.R")

url2 <- "https://raw.githubusercontent.com/cindyellow/SURA-2022/main/preprocess_smoke_kml.R"
if (!file.exists("../preprocess_smoke_kml.R"))
  download.file(url2, destfile = "../preprocess_smoke_kml.R")
source("../preprocess_smoke_kml.R")
```


## HDBSCAN: Assigning representative points for each cluster

```{r}
# Function for finding best minpts for a day and building clusters based on that
build_best_cl <- function(day){
  coords <- as.tibble(st_coordinates(day$geometry))
  scores <- c()
  for (x in minpts){
    cl <- hdbscan(coords, minPts = x)
    # check if there are cluster scores available
    if (all(is.na(cl$cluster_scores))){
      scores <- c(scores, 0)
    } else{
      scores <- c(scores, cdbw(coords, cl$cluster)$cdbw)
    }
  }
  
  best_param <- minpts[which.max(scores)]
  return (hdbscan(coords, minPts=best_param))
}
```

```{r}
# Function for getting k representative points
get_rep_pts <- function(cl, day, k){
  # Gives us highest member probability of each cluster
  mem_probs <- data.frame(cluster = cl$cluster, prob = cl$membership_prob)
  mem_probs <- mem_probs %>%
    group_by(cluster) %>%
    summarise(max_prob = max(prob))
  # Assign cluster and member probability to all observations 
  day <- day %>%
    mutate(cluster = cl$cluster, mem_prob = cl$membership_prob) %>%
    filter(cluster != 0)
  # Group by cluster
  cluster_info <- as.tibble(day) %>%
    group_by(cluster) %>%
    summarise(frp_avg = round(mean(frp, na.rm=TRUE),4),
              frp_vars = round(var(frp, na.rm=TRUE),4),
              num_pts = n()) 
  reps <- day %>%
    inner_join(cluster_info, by="cluster") %>%
    mutate(cluster = paste0(date, "_", cluster)) %>%
    arrange(desc(mem_prob)) %>%
    group_by(cluster) %>%
    distinct(geometry, .keep_all=TRUE) %>%
    slice(1:k)  
  return(reps)
}
```

```{r}
# Get representative points for all dates within range
minpts <- seq(5,100,by=5)
rep_pts <- c()
dates <- seq(as.Date("2018-11-08"), as.Date("2018-11-16"), by=1)

for (d in as.list(dates)){
  day <- in_cali %>%
    filter(date == d)
  cl <- build_best_cl(day)
  reps <- get_rep_pts(cl, day, 1)
  rep_pts <- rbind(rep_pts, reps)
}
```

```{r}
head(rep_pts)
```

Since our end goal is to calculate the distance between other points of interest and the fire cluster, I decided to first filter by unique points for each cluster before assigning the top k representative points. This will ensure that we have a diverse representation for each cluster.

## Merging with AQS data
```{r}
# Read AQS data and restrict to campfire week
aqs <- read.delim("../data/AQS_PM25_2000_2021_Cali.csv", sep=",", strip.white=TRUE)
aqs <- aqs %>%
  filter(Date > "2018-11-06" & Date < "2018-11-17")
aqs <- aqs %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, remove=FALSE) %>%
  st_set_crs(4326)
```

```{r}
rep_pts_sf <- st_as_sf(rep_pts)
smoke <- st_as_sf(smoke)
```

### Calculate distance between each AQS point and nearest fire


## Indicator Variables

```{r}
#Filter by date for all dfs
dates <- seq(as.Date("2018-11-08"), as.Date("2018-11-16"), by=1)

# Set default values for smoke indicator variables
aqs <- aqs %>%
  mutate(light=NA, med=NA, heavy=NA,fire_dist=NA,closest_cl=NA)

for (d in as.list(dates)){
  day_smoke <- smoke %>%
    filter(format(start_time, format="%Y-%m-%d") == d)
  day_aqs <- aqs %>%
    dplyr::select(c("PM25", "Date", "Latitude", "Longitude")) %>%
    filter(Date == d)
  day_fire <- rep_pts_sf %>%
    filter(date == d)
  
  # Get closest cluster & distance
  closest_fire <- st_nearest_feature(day_aqs, day_fire)
  day_aqs$fire_dist <- st_distance(day_aqs$geometry, day_fire[closest_fire,], by_element=TRUE)
  day_aqs$closest_cl <- day_aqs$closest_cl <- day_fire[closest_fire,]$cluster
  
  # Get indicator variables
  smoke_regions <- st_intersects(day_aqs$geometry, day_smoke$geometry)
  day_aqs$light <- lapply(smoke_regions, function(x)ifelse('light' %in% unique(day_smoke[unlist(x),]$type), 1, 0))
  day_aqs$med <- lapply(smoke_regions, function(x)ifelse('medium' %in% unique(day_smoke[unlist(x),]$type), 1, 0))
  day_aqs$heavy <- lapply(smoke_regions, function(x)ifelse('heavy' %in% unique(day_smoke[unlist(x),]$type), 1, 0))
  day_aqs <- as.tibble(day_aqs) %>%
    dplyr::select(-geometry) %>%
    mutate(across(c("light", "med", "heavy"), as.double))
  
  aqs <- aqs %>%
    left_join(as.tibble(day_aqs), by=c("PM25", "Date", "Latitude", "Longitude")) %>%
    mutate(light = coalesce(light.y, light.x),
           med = coalesce(med.y, med.x),
           heavy = coalesce(heavy.y, heavy.x),
           fire_dist = coalesce(fire_dist.y, fire_dist.x),
           closest_cl = coalesce(closest_cl.y, closest_cl.x)) %>% 
    dplyr::select(-light.x, -light.y, -med.x, -med.y, -heavy.x, -heavy.y, -fire_dist.x, -fire_dist.y, -closest_cl.x, -closest_cl.y)
}

```

### Doublecheck

```{r}
#??? don't know what's going on
match <- aqs$Date == as.Date(gsub("_[0-9][0-9]*", "", aqs$closest_cl))
as.tibble(aqs[match==FALSE,])

# Drop null observations
aqs_na <- aqs %>%
  drop_na(Longitude)

sum(aqs$Longitude=="NA")

```


## Cluster Polygons
```{r}
# https://stackoverflow.com/questions/53440909/polygon-from-cluster-of-lat-long-points-in-r

# https://gis.stackexchange.com/questions/303524/clustering-spatially-connected-polygons-so-all-clusters-have-approximately-the-s
```



