---
title: "Week 1: Data Exploration"
subtitle: "SURA 2022"
author: "Shih-Ting (Cindy) Huang"
date: "06/01/2022"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo=FALSE)

library(tidyverse)
library(dplyr)
library(leaflet)
```

# Data Information
## Time Frame

- The HMS system has operated 24x7x365 since June 16th, 2003, when some of its first satellites have been put to use. Some of these satellite sources have been discontinued after some time, but Terra/MODIS is one of the first satellites used and is still in use. 
- However, I noted that in the first few years up until approixmately November 2016, all the fire pixel calculations were done manually, so we don't have an FRP score with me,

## Data Structure, Variables, Etc.

- They record fire detection data several times a day (around once every 2 hours), but there can be a little delay in making the data available (this doesn't affect the observation time recorded).

- The variables provided are:
  
  - Longitude
  - Latitude
  - Observation date (year & Julian day 0-365)
  - Observation time (in UTC)
  - Satellite/sensor used
  - Method of detection (either manual or automated)
  - Ecosystem type
  - Fire Radiative Power (FRP) in megawatts

In particular, FRP can be used to measure fire. 

## Data Sources

- Satellites used include GOES, Terra/MODIS, Aqua/MODIS, NOAA, MetOp, and S-NPP/VIIRs
- Even though most of the satellites only operated for a certain time, I think we can still use HMS as our main data source because Terra/MODIS has been present from the beginning. This can be used as a baseline, and we can draw data from other satellites to compare with it.


# Sample Data Visualization
## Data Cleaning & Wrangling

The steps I took were:

- Renamed columns and converted them to the right type
- Changed filler values for FRP (-999.000) to NA

```{r}
# Read the text file
may1 <- read.delim("https://raw.githubusercontent.com/cindyellow/SURA-2022/main/data/fire/hms20220501.txt", sep=",", strip.white=TRUE)
```

```{r, eval=FALSE}
str(may1)
```

```{r, eval=FALSE}
head(may1)
```

```{r}
# Rename columns
may1 <- may1 %>%
  rename(
    longitude = Lon,
    latitude = Lat,
    date = YearDay,
    time = Time, 
    satellite = Satellite,
    method_of_detect = Method.of.Detect,
    ecosys = Ecosys,
    frp = Fire.RadPower
  )
```

```{r}
# Convert to correct variable type
may1 <- may1 %>%
  mutate(
    date = as.Date(as.character(date),          
                   format = "%Y%j"),
    method_of_detect = as.factor(method_of_detect),
    satellite = as.factor(satellite),
    ecosys = as.factor(ecosys),
    time = substr(as.POSIXct(sprintf("%04.0f", time), format='%H%M'), 12, 16)
  )
```

```{r}
# Add a new column for date and time
may1 <- may1 %>%
  mutate(
    date_comp = as.POSIXct(paste(date, time), 
                            format = "%Y-%m-%d %H:%M")
  )
```


```{r}
# Convert fill values to NA for FRP
may1 <- may1 %>%
   mutate(frp = na_if(frp, -999.000))
```

```{r}
unique(may1$method_of_detect)
```


```{r}
# Take a look at the distribution of FRP
may1 %>%
  ggplot(aes(x=frp)) +
  geom_histogram(fill="#727D71") +
  theme_minimal() +
  xlab("FRP") + 
  ylab("Frequency") +
  labs(title = "Histogram for FRP") +
  theme(plot.title = element_text(size = 12, hjust = 0.5))
```

## Daily

As a sample data, we take the observations from one day (05/01/2022) to plot onto a map. Below is a simple leaflet map:

```{r}
library(leaflet.extras2)

det_pal <- colorFactor(c('#5E565A', '#8B80F9', '#A9CBB7', '#FF934F'), domain = may1$method_of_detect[!is.na(may1$method_of_detect)], na.color = NA)

labels <- paste(
  paste0("FRP: ", may1$frp),
 paste0("<br>Ecosys: ", may1$ecosys)) %>%
  lapply(htmltools::HTML)

may1 %>%
  leaflet() %>%
  addProviderTiles('OpenStreetMap') %>%
  addCircles(lat=~latitude, lng=~longitude, color=~det_pal(method_of_detect), 
             label =~labels, 
             opacity = 1, fillOpacity=1, radius=~30*frp) %>%
  addLegend('bottomleft', pal = det_pal, values = may1$method_of_detect, title = "Method of Detection", opacity=1, na.label=NULL)
```

Next, we show a map but with the ability to filter by ecosystem type.

```{r}
library(crosstalk)
library(tidyverse)
library(leaflet.extras)

may_sd<- SharedData$new(may1)

map<- leaflet(may_sd) %>%
  addProviderTiles('OpenStreetMap') %>%
  addCircles(lat=~latitude, lng=~longitude, color=~det_pal(method_of_detect), label =~paste('FRP', frp), opacity = 1, fillOpacity=1, radius=~30*frp) %>%
  addLegend('bottomleft', pal = det_pal, values = may1$method_of_detect, title = "Method of Detection", opacity=1, na.label=NULL)

#add filter
bscols(
filter_select("Ecosys", "Ecosys", may_sd, ~ecosys)
  )

bscols(map)
```

[WORK IN PROGRESS] Since the data is spatiotemporal, I'm trying to see a way to show the distribution of FRP overtime with a time slider. The visuals still aren't completely clear:

```{r}
library(leaflet.extras2)
library(sf)
library(geojsonsf)

projcrs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

data <- sf::st_as_sf(may1,
                     coords = c("longitude", "latitude"),
                     crs=projcrs)
data <- st_cast(data, "POINT")

leaflet() %>%
  addTiles() %>%
  addTimeslider(data=data,
             options = timesliderOptions(
               position = "topright",
               timeAttribute = "date_comp",
               range = TRUE)) %>%
  setView(-72, 22, 4)
```


```{r}
library(sf) #very useful!
smoke <- read_sf("data/smoke/smoke20220501.kml")
```

```{r}
smoke %>%
  glimpse()
```






