---
title: "Week 2: Data Exploration"
subtitle: "SURA 2022"
author: "Shih-Ting (Cindy) Huang"
date: '2022-06-08'
output: html_document
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo=FALSE)

library(tidyverse)
library(dplyr)
library(leaflet)
library(sf)
library(mapview)
library(rgdal)
library(shiny)
```

## HMS Fire Data: More Exploration

```{r}
fire <- read.delim("data/fire/hms20220501.txt", sep=",", strip.white=TRUE)
```

```{r}
# Rename columns
fire <- fire %>%
  rename(
    longitude = Lon,
    latitude = Lat,
    date = YearDay,
    time = Time, 
    satellite = Satellite,
    method_of_detect = Method.of.Detect,
    ecosys = Ecosys,
    frp = Fire.RadPower
  ) %>%
  mutate(
    date = as.Date(as.character(date),          
                   format = "%Y%j"),
    method_of_detect = as.factor(method_of_detect),
    satellite = as.factor(satellite),
    ecosys = as.factor(ecosys),
    time = substr(as.POSIXct(sprintf("%04.0f", time), format='%H%M'), 12, 16)
  )
```

```{r}
# Add a new column for date and time
fire <- fire %>%
  mutate(
    date_comp = as.POSIXct(paste(date, time), 
                            format = "%Y-%m-%d %H:%M"),
    frp = na_if(frp, -999.000)
  )
```

```{r}
# Combine data from different days
filelist = list.files(path="./data/fire/", pattern = "*.txt")

#assuming tab separated values with a header    
datalist = lapply(filelist, function(x)read.delim(paste("./data/fire/",x,sep=""), sep=",", strip.white=TRUE))

#assuming the same header/columns for all files
califire <- do.call("rbind", datalist) 

# Filter by California/West Coast
califire <- califire %>%
  filter(between(Lon,-125,-117) & between(Lat,32,44))

# Preprocess them
# Rename columns
califire <- califire %>%
  rename(
    longitude = Lon,
    latitude = Lat,
    date = YearDay,
    time = Time, 
    satellite = Satellite,
    method_of_detect = Method.of.Detect,
    ecosys = Ecosys,
    frp = Fire.RadPower
  ) %>%
  mutate(
    date = as.Date(as.character(date),          
                   format = "%Y%j"),
    method_of_detect = as.factor(method_of_detect),
    satellite = as.factor(satellite),
    ecosys = as.factor(ecosys),
    time = substr(as.POSIXct(sprintf("%04.0f", time), format='%H%M'), 12, 16)
  )
# Add a new column for date and time
califire <- califire %>%
  mutate(
    date_comp = as.POSIXct(paste(date, time), 
                            format = "%Y-%m-%d %H:%M"),
    frp = na_if(frp, -999.000)
  )
```

### Device 
There is correspondence between method of detection and satellites:

- VIIRs: SUOMI NPP, NOAA20
- FDC: GOES-EAST
- MODIS: MODIS-TERRA

Plotting another date, we see that the satellites don't have a regular record time. 

```{r, warning=FALSE}
# Fire detection time by device
fire %>%
  ggplot(aes(x=date_comp, fill=satellite)) +
  geom_histogram(bins=30) +
  xlim(as.POSIXct("2022-05-01 00:00"), as.POSIXct("2022-05-01 23:59"))+
  theme_minimal() +
  xlab("Record Time") + 
  ylab("Frequency") +
  labs(title = "Histogram for Time") +
  theme(plot.title = element_text(size = 12, hjust = 0.5))

fire %>%
  ggplot(aes(x=date_comp, fill=method_of_detect)) +
  geom_histogram(bins=30) +
  xlim(as.POSIXct("2022-05-01 00:00"), as.POSIXct("2022-05-01 23:59"))+
  theme_minimal() +
  xlab("Record Time") + 
  ylab("Frequency") +
  labs(title = "Histogram for Time") +
  theme(plot.title = element_text(size = 12, hjust = 0.5))
```
```{r}
califire %>%
  filter(format(as.Date(date_comp,format="%Y-%m-%d %H:%M"), format = "%d") == "08") %>%
  ggplot(aes(x=date_comp, fill=satellite)) +
  geom_histogram(bins=30) +
  xlim(as.POSIXct("2018-11-08 00:00"), as.POSIXct("2018-11-08 23:59"))+
  theme_minimal() +
  xlab("Record Time") + 
  ylab("Frequency") +
  labs(title = "Histogram for Time") +
  theme(plot.title = element_text(size = 12, hjust = 0.5))

```

### Ecosys

The different types of ecosystems are listed in the [GLCC README file](https://www.usgs.gov/media/files/global-land-cover-characteristics-data-base-readme-version2) under Appendix 1 "Global Ecosystems Legend". These categories have been created based on " 1-km AVHRR (Advanced Very High Resolution Radiometer) 10-day NDVI (Normalized Difference Vegetation Index ) composites". The ones we have in the dataset (during the Californian fire) are:

```{r}
califire %>%
  ggplot(aes(ecosys)) +
  geom_bar()+
  theme_minimal() +
  xlab("Ecosystem") + 
  ylab("Frequency") +
  labs(title = "Bar Graph for Ecosystem Code") +
  theme(plot.title = element_text(size = 12, hjust = 0.5))
```

From the bar chart, we can see that the following ecosystems are the most prominent during the Californian fire:

- 22: Cool Conifer Forest 
- 46: Mediterranean Scrub 
- 26: Deciduous Broadleaf Forest 
- 24: Mixed Forest 
- 91: Woody Savanna

### Fire Detection Overtime (Weeks)

```{r}
# Plot histogram by date
datelist <- unique(califire$date)

draw_plot <- function(date_to_filter_by) {
  filtered_cal <- califire %>%
    filter(date == !!date_to_filter_by)
  ggplot(filtered_cal, aes(x=date_comp, fill=satellite)) +
    geom_histogram(bins=30) +
    theme_minimal() +
    xlab("Record Time") + 
    ylab("Frequency") +
    labs(title = "Histogram for Time") +
    theme(plot.title = element_text(size = 12, hjust = 0.5))
}

ui <- fluidPage(
  inputPanel(
    selectInput(
      "PlotDate",
      label = "Select Date",
      choices = datelist
    )
  ),
  plotOutput('minplot')
)

server <- function(input, output) {
  output$minplot <- renderPlot(draw_plot(input$PlotDate))
}

shinyApp(ui = ui, server = server)
```


## Relating Smoke & Fire

```{r}
# https://stackoverflow.com/questions/50775357/how-to-read-in-kml-file-properly-in-r-or-separate-out-lumped-variables-into-col
smoke <- st_read("data/smoke/smoke20220501.kml", layer="Folder")
```

```{r}
# Calculate area
smoke$area <- st_area(smoke)

# Extract density
# TODO: figure out why all densities are the same, how to read in layers
smoke$density <- str_remove_all(str_extract(smoke$Description, "Density: [0-9]"),"[a-zA-Z]+: ")
```

```{r}
fire <- st_as_sf(fire, coords = c("longitude", "latitude"))
st_crs(fire) <- "+init=epsg:4326 +proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0" # assign CRS to points
```





