---
title: "Week 5-6: HDBSCAN Clustering" 
subtitle: "SURA 2022"
author: "Shih-Ting (Cindy) Huang"
date: '2022-07-06'
output: html_document
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo=FALSE)

library(tidyverse)
library(dplyr)
library(leaflet)
library(sf)
library(mapview)
library(ggplot2)
library(RColorBrewer)
library(dbscan)
library(leafpop)
```


```{r load-data, echo=FALSE, include=FALSE}
# Run the script to preprocess fire datasets to include information on time, location, and FRP
url1 <- "https://raw.githubusercontent.com/cindyellow/SURA-2022/main/preprocess_fire_kml.R"
if (!file.exists("../preprocess_fire_kml.R"))
  download.file(url1, destfile = "../preprocess_fire_kml.R")
source("../preprocess_fire_kml.R")

url2 <- "https://raw.githubusercontent.com/cindyellow/SURA-2022/main/preprocess_smoke_kml.R"
if (!file.exists("../preprocess_smoke_kml.R"))
  download.file(url2, destfile = "../preprocess_smoke_kml.R")
source("../preprocess_smoke_kml.R")
```


## HDBSCAN 

```{r}
# Try HDBSCAN for 11/08/2018
one <- in_cali %>%
  filter(date == "2018-11-08")

coords <- as.tibble(st_coordinates(one$geometry))
```


### Finding the best parameters

There are several metrics for evaluating the quality of clusters. The following are some common ones and an evaluation:

- CDbw: considers multiple representative points per cluster, but it means that we have to introduce a parameter for defining the number of points
  - cohesion: within-cluster density change
  - compactness: average within-cluster density
  - separation: inter-cluster density
  - CDbw index: relative validity index that takes into consideration all of the aforementioned metrics
- DBCV: evaluates clusters by comparing the highest density between clusters and the lowest density in clusters

Currently, CDbw is used to compare the clusters because there is a well-implemented package in R for it. 

```{r}
library(fpc)

# Compare CDBW index for all possible clusters
minpts <- seq(5,100, by=5)
scores <- c()
for (x in minpts){
  cl <- hdbscan(coords, minPts = x)
  scores <- c(scores, cdbw(coords, cl$cluster)$cdbw)
}

best_param <- minpts[which.max(scores)]
```

```{r}
# Create hdbscan object with best minPts
cl <- hdbscan(coords, minPts = best_param)
plot(coords, col=cl$cluster+1, pch=20)
```
Based on comparing the CDbw indices for each model, the best minPts is `r best_param` that creates `r length(unique(cl$cluster))` clusters.


## Cluster Visualization

In HDBSCAN, clusters are arbitrarily shape. Hence, we cannot obtain a cluster "centroid" effectively. Hence, we still display all the points colored by their clusters in the visualization.

```{r}
# Add cluster assignments to the original dataset
one <- one %>%
  mutate(cluster = cl$cluster) %>%
  filter(cluster != 0) # Eliminates noise points

one <- st_as_sf(one)
smoke <- st_as_sf(smoke)
```



```{r}
mapview(smoke, zcol ="type", label=FALSE, legend = TRUE, layer.name = 'Smoke Type',col.regions = brewer.pal(3, "RdYlGn"), col="transparent",
        popup = popupTable(
                smoke,
                zcol = c(
                  "start_time",
                  "end_time",
                  "density",
                  "area"
                  ),
                feature.id = FALSE
                )        
        )+
mapview(one, zcol = "cluster", label=FALSE, legend=FALSE, layer.name = 'cluster', 
        popup = popupTable(
              one,
              zcol = c(
                "start_time",
                "method_of_detect",
                "satellite",
                "cluster"
                ),
              feature.id = FALSE
              )
        )
```

## Cluster FRP: Summary

```{r}
# Group by cluster
cluster_info <- as.tibble(one) %>%
  group_by(cluster) %>%
  summarise(frp_avg = round(mean(frp, na.rm=TRUE),4),
            frp_vars = round(var(frp, na.rm=TRUE),4)) 
# Assign them with average
one <- as.tibble(one) %>%
  inner_join(cluster_info, by="cluster")
```

```{r}
# Display the average FRP for each cluster and the variance
tibble("Cluster" = cluster_info$cluster, "Average FRP" = cluster_info$frp_avg, "Variance" = cluster_info$frp_vars) %>%
  knitr::kable(caption="A summary of clusters for 11/08/2018")
```

Based on the cluster summary above, we can see that the average FRP was the highest in clusters 19, 7, and 2. Note that we can't compare them in an absolute sense since the same FRP value could have different implications depending on the ecosystem.

## PostGIS

- PostgreSQL but with spatial functionality
- [Resource for setting up](https://towardsdatascience.com/spatial-data-science-with-postgresql-postgis-2f941c8c367a)

## References: 

| 1. Moulavi, Davoud & A Jaskowiak, Pablo & Campello, Ricardo & Zimek, Arthur & Sander, Joerg. (2014). Density-Based Clustering Validation. 10.1137/1.9781611973440.96. 

|    [Access the paper here](https://www.researchgate.net/publication/260333211_Density-Based_Clustering_Validation)

| 2. [A density-based cluster validity approach using multi-representatives](https://www.sciencedirect.com/science/article/pii/S0167865508000020)


