---
title: "Week 7: HDBSCAN Clustering P3, Merging with AQS" 
subtitle: "SURA 2022"
author: "Shih-Ting (Cindy) Huang"
date: '2022-07-13'
output: html_document
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo=FALSE)

library(tidyverse)
library(dplyr)
library(leaflet)
library(sf)
library(mapview)
library(ggplot2)
library(RColorBrewer)
library(dbscan)
library(leafpop)
library(fpc)
```


```{r load-data, echo=FALSE, include=FALSE}
# Run the script to preprocess fire datasets to include information on time, location, and FRP
url1 <- "https://raw.githubusercontent.com/cindyellow/SURA-2022/main/preprocess_fire_kml.R"
if (!file.exists("../preprocess_fire_kml.R"))
  download.file(url1, destfile = "../preprocess_fire_kml.R")
source("../preprocess_fire_kml.R")

url2 <- "https://raw.githubusercontent.com/cindyellow/SURA-2022/main/preprocess_smoke_kml.R"
if (!file.exists("../preprocess_smoke_kml.R"))
  download.file(url2, destfile = "../preprocess_smoke_kml.R")
source("../preprocess_smoke_kml.R")
```


## HDBSCAN: Assigning representative points for each cluster

```{r}
# Function for finding best minpts for a day and building clusters based on that
build_best_cl <- function(day){
  coords <- as.tibble(st_coordinates(day$geometry))
  scores <- c()
  for (x in minpts){
    cl <- hdbscan(coords, minPts = x)
    # check if there are cluster scores available
    if (is.na(cl$cluster_scores)){
      scores <- c(scores, 0)
    } else{
      scores <- c(scores, cdbw(coords, cl$cluster)$cdbw)
    }
  }
  
  best_param <- minpts[which.max(scores)]
  return (hdbscan(coords, minPts=best_param))
}
```

```{r}
# Function for getting k representative points
get_rep_pts <- function(cl, day, k){
  # Gives us highest member probability of each cluster
  mem_probs <- data.frame(cluster = cl$cluster, prob = cl$membership_prob)
  mem_probs <- mem_probs %>%
    group_by(cluster) %>%
    summarise(max_prob = max(prob))
  # Assign cluster and member probability to all observations 
  day <- day %>%
    mutate(cluster = cl$cluster, mem_prob = cl$membership_prob) %>%
    filter(cluster != 0)
  # Group by cluster
  cluster_info <- as.tibble(day) %>%
    group_by(cluster) %>%
    summarise(frp_avg = round(mean(frp, na.rm=TRUE),4),
              frp_vars = round(var(frp, na.rm=TRUE),4),
              num_pts = n()) 
  reps <- day %>%
    inner_join(cluster_info, by="cluster") %>%
    arrange(desc(mem_prob)) %>%
    group_by(cluster) %>%
    distinct(geometry, .keep_all=TRUE) %>%
    slice(1:k)  
  return(reps)
}
```


```{r}
# Get representative points for all dates within range
minpts <- seq(5,100,by=5)
rep_pts <- c()
dates <- seq(as.Date("2018-11-07"), as.Date("2018-11-16"), by=1)

for (d in as.list(dates)){
  day <- in_cali %>%
    filter(date == d)
  cl <- build_best_cl(day)
  reps <- get_rep_pts(cl, day, 1)
  rep_pts <- rbind(rep_pts, reps)
}
```

```{r}
head(rep_pts)
```

Since our end goal is to calculate the distance between other points of interest and the fire cluster, I decided to first filter by unique points for each cluster before assigning the top k representative points. This will ensure that we have a diverse representation for each cluster. However, if we want to consider the intensity/duration of the fire at that location, we might consider not doing this? **

## Merging with AQS data
```{r}
# Read AQS data and restrict to campfire week
aqs <- read.delim("../data/AQS_PM25_2000_2021_Cali.csv", sep=",", strip.white=TRUE)
aqs <- aqs %>%
  filter(Date > "2018-11-07" & Date < "2018-11-17")
```

```{r}
rep_pts_sf <- st_as_sf(rep_pts)
smoke <- st_as_sf(smoke)
aqs <- aqs %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, remove=FALSE)
```

## Visualization

```{r}

m <- mapview(rep_pts_sf, zcol = "date", label=FALSE, legend=FALSE, layer.name=c("Fire Points"),col.regions = brewer.pal(10, "Spectral"),
        popup = popupTable(
              rep_pts_sf,
              zcol = c(
                "date",
                "method_of_detect",
                "satellite",
                "cluster"
                ),
              feature.id = FALSE
              )
        ) + 
  mapview(smoke, zcol ="type", label=FALSE, legend = TRUE, layer.name = 'Smoke Type',col.regions = brewer.pal(3, "RdYlGn"), col="transparent",
        popup = popupTable(
                smoke,
                zcol = c(
                  "start_time",
                  "end_time",
                  "density",
                  "area"
                  ),
                feature.id = FALSE
                )        
        )

# Add AQS points as markers

my_icons <- iconList(star <- makeIcon(iconUrl = "https://img.icons8.com/windows/32/383838/marker.png",
                          iconWidth = 25, iconHeight = 25))
m <- m@map %>%
  addMarkers(~Longitude, ~Latitude,
             icon=~my_icons,
             popup=~paste(paste0("Date: ", Date),
                        paste0("PM 2.5: ", PM25),
                      sep="<br>"),
             data=aqs)

m
```

```{r}
# include aqs distance to cluster? maybe take min distance 
```












