---
title: "Week 4: Fire Data Prepreprocessing & HDBSCAN"
subtitle: "SURA 2022"
author: "Shih-Ting (Cindy) Huang"
date: '2022-06-22'
output: html_document
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)

library(tidyverse)
library(dplyr)
library(leaflet)
library(sf)
library(mapview)
library(rgdal)
library(shiny)
library(ggplot2)
library(rvest)
library(RColorBrewer)
library(leafpop)
```

## Fire Data: Preprocessing

### Combining all California Campfire Data

In this section, we read and process all the fire data for the California Campfire week. 
```{r, echo=T, results='hide'}
# Combine data from different days
filelist = list.files(path="../data/fire/", pattern = "*.kml")

#assuming tab separated values with a header    
datalist = lapply(filelist, function(x)st_read(paste("../data/fire/", x, sep="")))

#assuming the same header/columns for all files
fire <- do.call("rbind", datalist) 
```

```{r, echo=T, results='hide'}
# Combine data from different days
filelist = list.files(path="../data/fire/", pattern = "*.txt")

#assuming tab separated values with a header    
datalist = lapply(filelist, function(x)read.delim(paste("../data/fire/",x,sep=""), sep=",", strip.white=TRUE))

#assuming the same header/columns for all files
firetxt <- do.call("rbind", datalist) 

firetxt <- firetxt %>%
  rename(
    longitude = Lon,
    latitude = Lat,
    date = YearDay,
    time = Time, 
    satellite = Satellite,
    method_of_detect = Method.of.Detect,
    ecosys = Ecosys,
    frp = Fire.RadPower
  ) %>%
  mutate(
    date = as.Date(as.character(date),          
                   format = "%Y%j"),
    method_of_detect = as.factor(method_of_detect),
    satellite = as.factor(satellite),
    ecosys = as.factor(ecosys),
    time = substr(as.POSIXct(sprintf("%04.0f", time), format='%H%M'), 12, 16)
  )
# Add a new column for date and time
firetxt <- firetxt %>%
  mutate(
    date_comp = as.POSIXct(paste(date, time), 
                            format = "%Y-%m-%d %H:%M"),
    frp = na_if(frp, -999.000)
  )
```

We also use the California boundary file to restrict our view to only points in California.

```{r, echo=T, results='hide'}
cal_bound <- st_read("../ca-state-boundary/CA_State_TIGER2016.shp")

# Convert to the same coordinate system as HMS (4326)
cal_bound <- cal_bound %>%
  st_set_crs(3857) %>% 
  st_transform(4326)
```

```{r}
# Get an array of whether each observation is in California
in_bound <- lengths(st_intersects(fire, cal_bound))>0
```

```{r}
# Get subset with indexing
# https://gis.stackexchange.com/questions/394954/r-using-st-intersects-to-classify-points-inside-outside-and-within-a-buffer
in_cali <- fire[in_bound,]
```

### Cleaning
```{r}
in_cali <- as.data.frame(in_cali) %>%
  dplyr::select(-Name) %>%
  # Remove prefix
  mutate(Description = gsub('Fire Attributes: YearDay: ', '', Description)) %>%
  # Replace column name strings with comma
  mutate(Description = gsub('[a-zA-Z]*: ', ',', Description)) %>%
  # Separate based on comma
  separate(., col=Description, 
           into = c('date', 'time', 'satellite','method_of_detect'),
           sep = ',') 
```

```{r}
in_cali <- in_cali %>% 
  # Remove unnecessary substrings
  mutate(satellite = (str_remove_all(satellite,"Method of"))) %>%
  # Convert into correct datetime
  mutate( time = substr(as.POSIXct(sprintf("%04.0f", as.integer(time)), format='%H%M'), 12, 16),
          start_time = as.POSIXct(paste(date, time),
                            format = "%Y%j %H:%M"),
          date = as.POSIXct(date, format = "%Y%j")) %>%
  dplyr::select(-time)
```

## Smoke Data: Preprocessing & Visualization

```{r, results='hide'}
# Read light, medium, heavy separately
# Combine data from different days
filelist = list.files(path="../data/smoke/", pattern = "*.kml")
filelist <- filelist[-1]

#assuming tab separated values with a header    
datalist = lapply(filelist, function(x)st_read(paste("../data/smoke/", x, sep=""), layer="Smoke (Light)"))

#assuming the same header/columns for all files
smoke_light <- do.call("rbind", datalist) 

# smoke_light <- st_read("../data/smoke/smoke20181108.kml", layer="Smoke (Light)")
smoke_light <- as.data.frame(smoke_light) %>%
  mutate(type="light")
```

```{r, results='hide'}
#assuming tab separated values with a header    
datalist = lapply(filelist, function(x)st_read(paste("../data/smoke/", x, sep=""), layer="Smoke (Medium)"))
smoke_med <- do.call("rbind", datalist) 
# smoke_med <- st_read("../data/smoke/smoke20181108.kml", layer="Smoke (Medium)")
smoke_med <- as.data.frame(smoke_med) %>%
  mutate(type="medium")
```

```{r, results='hide'}
datalist = lapply(filelist, function(x)st_read(paste("../data/smoke/", x, sep=""), layer="Smoke (Heavy)"))
smoke_heavy <- do.call("rbind", datalist) 
# smoke_heavy <- st_read("../data/smoke/smoke20181108.kml", layer="Smoke (Heavy)")
smoke_heavy <- as.data.frame(smoke_heavy) %>%
  mutate(type="heavy")

smoke <- list(smoke_light, smoke_med, smoke_heavy)

smoke <- smoke %>%
  reduce(full_join, by=c('Name', 'Description', 'type','geometry'))
```

```{r, results='hide'}
smoke <- smoke %>%
  dplyr::select(-Name) %>%
  mutate(Description = gsub('Smoke Attributes: Start Time: ', '', Description)) %>%
  mutate(Description = gsub('[a-zA-Z]*: ', ',', Description)) %>%
  separate(., col=Description, 
           into = c('start_time', 'end_time', 'density', 'satellite'),
           sep = ',') 
```

```{r, results='hide'}
# Switch off spherical geometry
sf_use_s2(FALSE)
# Clean time and add smoke area
smoke <- smoke %>% 
  # Remove unnecessary substrings
  mutate(start_time = (str_remove_all(start_time,"[a-zA-Z]")),
         end_time = (str_remove_all(end_time,"[a-zA-Z]"))) %>%
  # Separate date and time based on space
  separate(., col=start_time, 
           into=c('start_date', 'st'),
           sep=' ') %>%
  separate(., col=end_time, 
           into=c('end_date','et'),
           sep=' ') %>%
  # Convert into datetime object
  mutate( st = substr(as.POSIXct(sprintf("%04.0f", as.integer(st)), format='%H%M'), 12, 16),
          et = substr(as.POSIXct(sprintf("%04.0f", as.integer(et)), format='%H%M'), 12, 16),
          start_time = as.POSIXct(paste(start_date, st),
                            format = "%Y%j %H:%M"),
         end_time = as.POSIXct(paste(end_date, et),
                            format = "%Y%j %H:%M"),
         area = st_area(smoke$geometry)) %>%
  dplyr::select(-start_date, -end_date, -st, -et)
```

## Mapping Data for the Week
```{r, echo=FALSE}
# Convert dataframes to sf objects for mapview
in_cali <- st_as_sf(in_cali)
```

```{r, echo=FALSE}
mapview(in_cali, zcol = "date", label=FALSE, legend=FALSE, layer.name = 'date', col.regions = brewer.pal(10, "Paired"),
        popup = popupTable(
              in_cali,
              zcol = c(
                "start_time",
                "method_of_detect",
                "satellite"
                ),
              feature.id = FALSE
              )
        )
```

## HDBSCAN 

First, we plot all the fire detection points in California.
```{r, echo=FALSE}
# Plot the points
library("dbscan")

plot(in_cali$geometry, pch=20)
```

Then, we plot it making adjustments to the coordinate system.

```{r, echo=FALSE}
# Convert to tibble object with separate columns for longitude and latitude
coords <- as.tibble(st_coordinates(in_cali$geometry))

# Plot the points again but now with coordinate system
plot(coords, pch=20)
```

Using these coordinates, we input into HDBSCAN specifying `minPts = 100`.
```{r, include=FALSE}
# Create hdbscan object 
cl <- hdbscan(coords, minPts = 100)
cl
```

We can plot the resulting clusters, excluding the outliers (belonging to `cluster = 0`).
```{r, echo=FALSE}
# Plot the clusters
plot(coords, col=cl$cluster+1, pch=20)
```

The points can also be shaded based on their membership probability. The higher the probability, the more solid the color.
```{r, echo=FALSE}
# Plot clusters and color by certainty of being in that cluster
plot(coords, col=cl$cluster+1, pch=21)
colors <- mapply(function(col, i) adjustcolor(col, alpha.f = cl$membership_prob[i]), 
                 palette()[cl$cluster+1], seq_along(cl$cluster))
points(coords, col=colors, pch=20)
```


